Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-08-30 pnas/Towards.tex
Updated 3-D experiments and cryo derivations

For the experiment in Figure~TKTK, we computed the analytical first three autocorrelations of the volume according to~TKTK. To estimate the coefficients of the volume itself, we solve the LS problem
\begin{equation} \label{eq:LS_cryo}
\min_{\hat{x},\hat{\gamma}} w_1 \vert \mathfrak{a}_x^1 - \hat{\gamma}\mathfrak{a}_{\hat{x}}^1 \vert^2 + w_2\| \mathfrak{a}_x^2 - \hat{\gamma}\mathfrak{a}_{\hat{x}}^2 \|_2^2 + w_3\| \mathfrak{a}_x^3 - \hat{\gamma}\mathfrak{a}_{\hat{x}}^3 \|_F^2, 
\end{equation}  
where the explicit expressions of $\mathfrak{a}_x^1$, $\mathfrak{a}_x^2$ and $\mathfrak{a}_x^3$ are given in~\eqref{eq:ac_1_prolates},~\eqref{eq:ac_2_prolates} and~\eqref{eq:ac_3_prolates}, respectively.  
In the experiments, we set $w_1=w_2=w_3=1$.

The LS was solved by TKTK. The expected $\hat \gamma$ is one.

In the first experiment, we used \TODO{Bla bla bla, real valued volume}

The micrographs for the experiment presented in Figure~\ref{fig:cryo_detection} were generated as follows \TODO{separation, etc. }. Then, the LS~\eqref{eq:LS_cryo} was solved with $L=0$ (low-resolution estimate).

@@

Let $\phi$ be the Coulomb potential representing the molecule we aim to recover. 
We assume that molecule is real-valued and smooth. In spherical coordinates, its 3-D Fourier transform $\widehat\phi$ admits a finite expansion of the form
\be\label{eq:volume_expansion} 
\widehat \phi(k, \theta, \varphi) = \sum_{\ell = 0}^L\sum_{m=-\ell}^{\ell}\sum_{s=1}^{S(\ell)}\tamir_{\ell, m, s}Y_{\ell}^m(\theta,\varphi)j_{\ell,s}(k),
\ee
where $Y_{\ell}^m$ are complex spherical harmonics and $j_{\ell,s}$ are normalized spherical Bessel functions; see Appendix TKTK for further definitions.  Let $I_{\omega}$ denote the tomographic projection obtained from viewing direction $\omega\in SO(3)$. By the Fourier projection-slice theorem, its 2-D Fourier transform  is given by~\cite{natterer1986mathematics}:
\be\label{eq:projection_model}
\widehat I_{\omega}(k,\varphi) = \sum_{\ell,m,m',s}\tamir_{\ell,m,s}D_{m',m}^{\ell}(\omega)Y_{\ell}^{m'}\left(\frac{\pi}{2},\varphi\right)j_{\ell,s}(k),\ee
where $D_{m',m}^{\ell}(\omega)$ is a Wigner-D matrix.

@@

Define the $p$th autocorrelation of $\II$ as
\begin{equation*} \label{eq:Kth_autocorrelation}
a^p_\II(\Bell_1,\ldots, \Bell_{p-1}) = \frac{1}{N^2}\sum_{\mb i}\II[\Bell]\II([\mb i+\Bell]_1)\cdots \II([\mb i+\Bell]_{p-1}),
\end{equation*}
where the summation is for $\mb i $ ranging over the $N^2$  pixels of the micrograph. %, $\Bell = (\Delta i, \Delta j)$.% and we set $\II(i,j)=0$ if either $i,j\notin\{1,\ldots, M\}$. 
%In our procedure, we compute the first three autocorrelations. %, called the mean, power spectrum, and bispectrum, of the micrographs. \TODO{This definition and terminology should come earlier}. 
Let $\II_1,\ldots\II_K$ denote a set of $K$ micrographs. 
Under the specified conditions, we show in Appendix~\ref{sec:moment_derivation} that the first three autocorrelations  of the micrographs are related to those of the projections by 
\begin{align} \label{eq:ac_micrographs}
\lim_{K\to\infty}\frac{1}{K}\sum_{i=1}^Ka^p_{\II_i}[\Bell_1,\ldots,\Bell_{p-1}]  &= \gamma\left\langle a^p_{I_\omega}[\Bell_1,\ldots,\Bell_{p-1}]\right\rangle_{\omega} \\&+ b_p[\Bell_1,\ldots,\Bell_{p-1}], \\ p=1,2,3, \quad \Bell_1,\ldots,\Bell_{p-1}&\in[-(P-1),P-1]^2, \nonumber
\end{align}
where $\langle\cdot\rangle_\omega$ denotes averaging over all possible viewing directions $\omega$ and $b_p$ is a bias term. 
Specifically,  $b_1 = 0$ and therefore the mean is unbiased. The bias term of the second-order autocorrelation  $b_2$ depends only on $\sigma^2$, the variance of the noise. Hence, if the noise level can be accurately estimated from the micrographs, this bias can be removed. 
Finally, the bias term of the third-order autocorrelation $b_3$ depends on the mean of the micrograph and $\sigma^2$.  Therefore, given sufficiently many projections, we can accurately estimate the quantities $\gamma\langle a^p_{I_{\omega}}\rangle_{\omega}$ directly from the micrographs. These quantities are functions of the unknown coefficients $\tamir_{l,m,s}$ and we could proceed to invert their relation, as we did in the 2-D toy example.

@@

In this section we prove relation~\eqref{eq:ac_micrographs}. We note that mathematically taking infinitely many micrographs is equivalent to taking one infinitely large micrograph with fixed density $\gamma$~\eqref{eq:gamma}. Hence, we consider the moments of one micrograph $\II$ in the limit $N\to\infty$ and  $\gamma = \lim_{N\to\infty}\frac{MP^2}{N^2}\in(0,1)$. 
The separation condition guarantees that if $\mb i=(i,j)$ is in the support of some projection, then $\mb i + \mb \Bell$ for $\Bell\in[-(P-1),P-1]^2$ is either in the support of the same projection or outside the support of any projection.

@@

We represent our autocorrelations using Prolate Spheroidal Wave Functions (PSWFs) $\{\psi_{k,q}\}$ where $k \geq 0, q \geq 1$ are integers~\cite{slepian1964pswfIV}. As we demonstrate below, this makes it easier to account for the fact that all in-plane rotations of the micrographs are equally likely observations. This is only a concern for the second and third order autocorrelations. Below, we start with $p=2$.
The PSWFs are eigenfunctions of the truncated Fourier transform and are given in polar coordinates by
\begin{align}
\psi_{k,q}(r,\varphi) & = \left\{\begin{array}{ll} \frac{1}{\sqrt{8\pi}}\alpha_{k,q}R_{k,q}(r)e^{\iota k\varphi}, & r\leq 1,\\ 0, & r>1,\end{array}\right. \label{eq:prolatesdef}
\end{align}
\TODO{Ref eq. 8 in Boris' paper}
where the ${R_{k,q}}$ are a family of real, one-dimensional functions and the ${\alpha_{k,q}}$ are  scaling factors which will be defined in the next section.
The PSWFs are orthogonal on the unit disk.

For $\Bell \in [-(P-1), P-1]^2$, let us define 
\begin{equation} \label{eq:PSWF_expansion}
\mathfrak{a}^2[k, q]  =
\sum_{\Bell}a^2_\II[\Bell]\overline{\psi_{k,q}[\Bell]}, \end{equation}
where $\psi_{k,q}[\Bell] := \psi_{k,q}(\Bell/(P-1))$ is a discretization of the PSWFs. \TODO{Range of $k,q$? need approx orthonormality in discrete...}
Knowledge of these coefficients is essentially equivalent to knowledge of the second-order correlations owing to the following approximate identity:
\begin{align}
a^2_\II[\Bell] & \approx \sum_{k,q}\mathfrak{a}^2[k,q]\psi_{k,q}[\Bell].
\label{eq:approxprolatesdiscrete}
\end{align}
This holds because the continuous PSWFs form an orthogonal basis, and their discretized counterparts are (empirically) almost orthogonal. As a result, for our purposes, the pair of equations above provides a basis expansion for the autocorrelations.

@@

Finally, we treat the bias terms, using approximate orthonormality \TODO{!} of the discretized PSFWs:
\begin{align*}
&\mk b_3[k_1,q_1; k_2,q_2] = \gamma\sigma^2\langle a^1_{I_{\omega}} \rangle_\omega \delta[k_1 - k_2]\\&\times \left[\delta[q_1-q_2] + \delta[k_1]\frac{1}{2\pi}(\alpha_{0,q_1} + \alpha_{0,q_2})R_{0,q_1}(0)R_{0,q_2}(0)\right].
\end{align*}
%
Thus,
\begin{align*}
b_3[\Bell_1, \Bell_2] & = \gamma\sigma^2\langle a^1_{I_{\omega}}\rangle_{\omega}\Big(\delta[\Bell_1 - \Bell_2] +
\delta[\Bell_1] + \delta[\Bell_2]\Big) \\
& \approx \sum_{\substack{k_1, q_1\\ k_2, q_2}} \mk 
b_3[k_1,q_1; k_2, q_2]\psi_{k_1,q_1}[\Bell_1]\overline{\psi_{k_2,q_2}[\Bell_2]}.
\end{align*}

\subsection{Connection to volume}

@@

We start by some preliminary computations. Let us  expand the Fourier transform of the volume as
\begin{equation} \label{eq:volume_expansion_app} 
\widehat \phi(ck, \theta, \varphi) = \sum_{\ell = 0}^L\sum_{m=-\ell}^{\ell}\sum_{s=1}^{S(\ell)}\tamir_{\ell, m, s}Y_{\ell}^m(\theta,\varphi)j_{\ell,s}(k), \quad k\leq 1,
\end{equation}
%\[ \widehat \phi(c\kk) = \sum_{\ell,m,s}a_{\ell,m,s}j_{\ell,s}(k)Y_{\ell,m}(\kk/k),\quad k\leq 1,\]
%where $k=||\kk||_2$, 
\TODO{How do we choose $S(\ell)$?}where $c$ is the bandlimit, $j_{\ell,s}$ is the normalized spherical Bessel functions given by
\[ j_{\ell, s}(k) = \frac{4}{|j_{\ell+1}(u_{\ell, s})|}j_{\ell}(2u_{\ell,s} k),\]
$j_{\ell}$ is the spherical Bessel function of order $\ell$ and  $u_{\ell,s}$ is the $s$th positive zero of $j_{\ell}$. We use 
the complex spherical harmonics $Y_{\ell}^m$ defined by  
\[ Y_{\ell,m}(\theta,\varphi) := \sqrt{\frac{2\ell+1}{4\pi}\cdot\frac{(\ell-m)!}{(\ell+m)!}}P_{\ell}^m(\cos\theta)e^{i m\varphi},\]
where $P_{\ell}^m$ are the associated Legendre polynomials with the Condon-Shortley phase. 
Sampling at the Nyquist-rate dictates $k\in[0,1/2]$~\cite{levin20173d}. \TODO{Didn't we say that $k\leq 1$?} With this notation, 
the Fourier transform of a  projection taken from the viewing direction $\omega\in SO(3)$ can be written as 
\[ \widehat I_{\omega}(ck,\theta) = \sum_{\ell,m,m',s}\tamir_{\ell,m,s}Y_{\ell,m'}(\pi/2,0)D_{m',m}^{\ell}(\omega)j_{\ell,s}(k)e^{im'\theta}.\]
%whenever $k\leq 1$. 
%where $D_{m',m}^{\ell}$ is the Wigner D-matrix.

Recall that the PSWFs are the solution of the integral equation \TODO{To mention the range of $\mb r$?} 
\begin{equation}
\label{eq:PSWF_defn_eq}
\alpha_{N,n}\psi_{N,n}(\mb k) = \int_{\mb r} \psi_{N,n}(\mb r)e^{\iota c(\mb r\cdot\mb k)}d\mb r,
\end{equation}
%\begin{equation}
%\label{eq:PSWF_defn_eq}
%\alpha_{k,q}\psi_{k,q}(\mb k) = \sum_{\mb i}\psi_{k,q}(\mb i)e^{-2\pi \iota (\mb i\cdot\mb k)}
%\end{equation}
and are given explicitly by \TODO{Do we use this normalization?}
\[ \psi_{N,n}(k,\theta) = \frac{1}{\sqrt{2\pi}}R_{N,n}(k)e^{\iota N\theta},\]
where $R_{N,n}(k)$ is a solution to another integral equation~\cite{landa2017steerable}. 
Let us assume that $I_{\omega}$ has bandlimit $c$ and is concentrated in the unit ball in real space. 
Using the 2-D PSWFs, we can express each projection as 
%Express this in 2D PSWFs so %\TODO{Eitan: I stopped here}
\[ \widehat I_{\omega}(ck,\theta) = \sum_{N,n}b_{N,n}(\omega)\psi_{N,n}(k,\theta),\]
%where in papers the 2D PSWFs are defined as
%\[ \psi_{N,n}(k,\theta) = \frac{1}{\sqrt{2\pi}}R_{N,n}(k)e^{ik\theta},\]
%%but in Boris' code the expansion is actually performed with respect to
%\[ \widetilde\psi_{N,n}(k,\theta) = \frac{1}{2\sqrt{2\pi}}\alpha_{N,n}R_{N,n}(k)e^{ik\theta},\]
%where $\alpha_{N,n}$ are the eigenvalues associated with $\psi_{N,n}$ - see Sect.~\ref{sec:PSWFs_props} below. 
%
%We then get
where
\[\begin{aligned} b_{N,n}(\omega) &= \frac{1}{\sqrt{2\pi}}\int_0^{2\pi}\int_0^1\widehat I_{\omega}(ck,\theta)R_{N,n}(k)e^{-\iota N\theta}k\, dk\, d\theta,\\
&= \sum_{\ell,m,m',s}\tamir_{\ell,m,s}\left[\sqrt{2\pi}Y_{\ell,m'}(\pi/2,0)\right]D_{m',m}^{\ell}(\omega)\\&\times \left(\int_0^1j_{\ell,s}(k)R_{N,n}(k)k\, dk\right)\left(\frac{1}{2\pi}\int_0^{2\pi}e^{\iota(m'-N)\theta}\, d\theta\right),\\
&= \sum_{\vert N\vert\leq \ell}\sum_{m,s}\tamir_{\ell,m,s}D_{N,m}^{\ell}(\omega)\beta_{\ell,s;N,n},\end{aligned}\]
where the coefficients
\begin{equation} \label{eq:beta}
\beta_{\ell,s;N,n} := \left\{\begin{array}{ll} \sqrt{2\pi}Y_{\ell,N}(\pi/2,0)\int_0^1j_{\ell,s}(k)R_{N,n}(k)k\, dk, & \ell\geq |N|\\ 0, & \ell<|N|\end{array}\right.,
\end{equation}
can be precomputed.
Since the PSWFs are eigenfunctions of the truncated Fourier operator~\eqref{eq:PSWF_defn_eq}, we can now express the projection in real space as
\begin{align} \label{eq:proj_expansion}
I_{\omega}(r,\varphi) &= \sum_{N,n}\widehat{\alpha}_{N,n}b_{N,n}(\omega)\psi_{N,n}(r,\varphi)\nonumber \\
&= \sum_{\ell=0}^L\sum_{N,m=-\ell}^{\ell}\sum_{n=0}^{n_{\text{max}}(N)}\sum_{s=1}^{S(\ell)}\tamir_{\ell,m,s}\widehat\beta_{\ell,s;N,n}D_{N,m}^{\ell}(\omega)\psi_{N,n}(r,\varphi)
\end{align}
\TODO{How do we choose $n_{max}$?}where $\alpha_{N,n}$ is the eigenvalue corresponding to the $(N,n)$th PSWF, $\widehat{\alpha}_{N,n} = (c/2\pi)^2\alpha_{N,n}$, and $\widehat\beta_{\ell,s;N,n}=\widehat\alpha_{N,n}\beta_{\ell,s;N,n}$.

@@

%Since the Fourier transformed volume is given by
%\[ \widehat V(k,\theta,\varphi) =
%\sum_{\ell,m,s}a_{\ell,m,s}Y_{\ell,m}(\theta,\varphi)j_{\ell,s}(k),\]
Since $j_{\ell,s}(0) = 0$ unless $\ell=0$, and since
$Y_{0,0}(\theta,\varphi) = \frac{1}{\sqrt{4\pi}}$, we conclude from~\eqref{eq:volume_expansion_app} that
\begin{equation} \label{eq:ac_1_prolates}
\mathfrak{a}_x^1 = \langle
a^1_{I_{\omega}}\rangle_{\omega}  = \widehat  \phi(\mb 0) = \frac{1}{\sqrt{4\pi}}\sum_s\tamir_{0,0,s}j_{0,s}(0).
\end{equation}

>>>>

\TODO{Fill in refs to eqns.} For the experiment in Figure~TKTK, we computed the analytical first three autocorrelations of the volume according to~TKTK. To estimate the coefficients of the volume itself, we solve the LS problem
\begin{equation} \label{eq:LS_cryo}
\min_{\hat{x},\hat{\gamma}} w_1 \vert \mathfrak{a}_x^1 - \hat{\gamma}\mathfrak{a}_{\hat{x}}^1 \vert^2 + w_2\| \mathfrak{a}_x^2 - \hat{\gamma}\mathfrak{a}_{\hat{x}}^2 \|_2^2 + w_3\| \mathfrak{a}_x^3 - \hat{\gamma}\mathfrak{a}_{\hat{x}}^3 \|_F^2, 
\end{equation}  
where the explicit expressions of $\mathfrak{a}_x^1$, $\mathfrak{a}_x^2$ and $\mathfrak{a}_x^3$ are given in~\eqref{eq:ac_1_prolates},~\eqref{eq:ac_2_prolates} and~\eqref{eq:ac_3_prolates}, respectively.  
In the experiments, we set $w_1=w_2=w_3=1$.

The LS problem was solved using Matlab's \texttt{lsqnonlin} solver for nonlinear least-squares problems. The expected $\hat \gamma$ in this experiment is one.

The true volume used in both experiments was the Bovine Pancreatic Trypsin Inhibitor (BPTI) mutant with altered binding loop sequence, whos atomic model is available in the Protein Data Bank (PDB)~\cite{pdb} as 1QLQ. We generated an EM map from this atomic model in UCSF Chimera~\cite{pettersen2004ucsf} at a resolution of 5 \AA, and cropped it to remove zeros at the boundary to obtain a volume of size $31^3$.

The micrographs for the experiment presented in Figure~\ref{fig:cryo_detection} were generated as follows. We sample rotation matrices from SO(3) uniformly at random using the QR-based algorithm described in QR, and generate the projection of the volume corresponding to that rotation matrix using \verb|cryo_project| in ASPIRE \cite{aspire}. We keep track of the indices at which the upper left corner of a projection can be placed without violating the separation condition, so all projections are separated by at least 30 pixels in each dimension. We continue adding projections to the micrograph until no more projections can be added without violating the separation condition. Then, the LS problem~\eqref{eq:LS_cryo} was solved assuming the spherical harmonic cutoff for the volume is $L=0$, which is sufficient to recover a significant $\widehat\gamma$ in the presence of projections in the micrograph.

@@

Let $\phi$ be the Coulomb potential representing the molecule we aim to recover. 
We assume that molecule is real-valued and smooth. In spherical coordinates, its 3-D Fourier transform $\widehat\phi$ admits a finite expansion of the form 
\begin{equation} \label{eq:volume_expansion_app} 
\widehat \phi(ck, \theta, \varphi) = \sum_{\ell = 0}^L\sum_{m=-\ell}^{\ell}\sum_{s=1}^{S(\ell)}\tamir_{\ell, m, s}Y_{\ell}^m(\theta,\varphi)j_{\ell,s}(k), \quad k\leq 1,
\end{equation}
%\[ \widehat \phi(c\kk) = \sum_{\ell,m,s}a_{\ell,m,s}j_{\ell,s}(k)Y_{\ell}^m(\kk/k),\quad k\leq 1,\]
%where $k=||\kk||_2$, 
where $c$ is the bandlimit, $\{S_l\}$ are determined using the Nyquist criterion as descirbed in \cite{bhamre2017anisotropic}, $j_{\ell,s}$ is the normalized spherical Bessel functions given by
\[ j_{\ell, s}(k) = \frac{4}{|j_{\ell+1}(u_{\ell, s})|}j_{\ell}(2u_{\ell,s} k),\]
$j_{\ell}$ is the spherical Bessel function of order $\ell$ and  $u_{\ell,s}$ is the $s$th positive zero of $j_{\ell}$. We use 
the complex spherical harmonics $Y_{\ell}^m$ defined by  
\[ Y_{\ell}^m(\theta,\varphi) := \sqrt{\frac{2\ell+1}{4\pi}\cdot\frac{(\ell-m)!}{(\ell+m)!}}P_{\ell}^m(\cos\theta)e^{i m\varphi},\]
where $P_{\ell}^m$ are the associated Legendre polynomials with the Condon-Shortley phase. 
Sampling at the Nyquist-rate dictates $c=1/2$~\cite{levin20173d}. 
Because $\phi$ is real-valued, so $\widehat \phi$ is conjugate-symmetric, the expansion coefficients satisfy $\tamir_{\ell, -m, s} = (-1)^{\ell+m}\overline{\tamir_{\ell,m,s}}$, and hence we only need to recover coefficients $\tamir_{\ell, m, s}$ with $m\geq 0$. Let $I_{\omega}$ denote the tomographic projection obtained from viewing direction $\omega\in SO(3)$. By the Fourier projection-slice theorem, its 2-D Fourier transform  is given by~\cite{natterer1986mathematics}:
\be\label{eq:projection_model}
\widehat I_{\omega}(k,\varphi) = \sum_{\ell,m,m',s}\tamir_{\ell,m,s}D_{m',m}^{\ell}(\omega)Y_{\ell}^{m'}\left(\frac{\pi}{2},\varphi\right)j_{\ell,s}(k),\ee
where $D_{m',m}^{\ell}(\omega)$ is a Wigner-D matrix. This implies that the projections are also $c$-bandlimited.

@@

Define the $p$th autocorrelation of $\II$ as
\begin{equation*} \label{eq:Kth_autocorrelation}
a^p_\II[\Bell_1,\ldots, \Bell_{p-1}] = \frac{1}{N^2}\sum_{\mb i}\II[\Bell]\II[\mb i+\Bell_1]\cdots \II[\mb i+\Bell_{p-1}],
\end{equation*}
where the summation is for $\mb i $ ranging over the $N^2$  pixels of the micrograph. %, $\Bell = (\Delta i, \Delta j)$.% and we set $\II(i,j)=0$ if either $i,j\notin\{1,\ldots, M\}$. 
%In our procedure, we compute the first three autocorrelations. %, called the mean, power spectrum, and bispectrum, of the micrographs. \TODO{This definition and terminology should come earlier}. 
Let $\II_1,\ldots\II_K$ denote a set of $K$ micrographs. 
Under the specified conditions, we show in Appendix~\ref{sec:moment_derivation} that the first three autocorrelations  of the micrographs are related to those of the projections by 
\begin{align} \label{eq:ac_micrographs}
\lim_{K\to\infty}\frac{1}{K}\sum_{i=1}^Ka^p_{\II_i}[\Bell_1,\ldots,\Bell_{p-1}]  &= \gamma\left\langle a^p_{I_\omega}[\Bell_1,\ldots,\Bell_{p-1}]\right\rangle_{\omega} \\&+ b_p[\Bell_1,\ldots,\Bell_{p-1}], \\ p=1,2,3, \quad \Bell_1,\ldots,\Bell_{p-1}&\in[-(P-1),P-1]^2, \nonumber
\end{align}
where $\langle\cdot\rangle_\omega$ denotes averaging over all possible viewing directions $\omega$ and $b_p$ is a bias term. 
Specifically,  $b_1 = 0$ and therefore the mean is unbiased. The bias term of the second-order autocorrelation  $b_2$ depends only on $\sigma^2$, the variance of the noise. Hence, if the noise level can be accurately estimated from the micrographs, this bias can be removed. 
Finally, the bias term of the third-order autocorrelation $b_3$ depends on the mean of the micrograph and $\sigma^2$.  Therefore, given sufficiently many projections, we can accurately estimate the quantities $\gamma\langle a^p_{I_{\omega}}\rangle_{\omega}$ directly from the micrographs. These quantities are functions of the unknown coefficients $\tamir_{l,m,s}$ and we could proceed to invert their relation, as we did in the 2-D toy example.

@@

In this section we prove relation~\eqref{eq:ac_micrographs}. We note that mathematically taking infinitely many micrographs is equivalent to take one infinitely large micrograph with fixed density $\gamma$. Hence, we consider the moments of one micrograph $\II$ in the limit $N\to\infty$ and  $\gamma = \lim_{N\to\infty}\frac{MP^2}{N^2}\in(0,1)$. 
The separation condition guarantees that if $\mb i=(i,j)$ is in the support of some projection, then $\mb i + \mb \Bell$ for $\Bell\in[-(P-1),P-1]^2$ is either in the support of the same projection or outside the support of any projection.

@@

We represent our autocorrelations using Prolate Spheroidal Wave Functions (PSWFs) $\{\psi_{k,q}\}$ where $k \geq 0, q \geq 1$ are integers~\cite{slepian1964pswfIV}. As we demonstrate below, this makes it easier to account for the fact that all in-plane rotations of the micrographs are equally likely observations. This is only a concern for the second and third order autocorrelations. Below, we start with $p=2$.
The PSWFs are eigenfunctions of the truncated Fourier transform and are given in polar coordinates by \footnote{A different normalization is used in \cite{landa2017steerable}.}
\begin{align}
\psi_{k,q}(r,\varphi) & = \left\{\begin{array}{ll} \frac{1}{\sqrt{8\pi}}\alpha_{k,q}R_{k,q}(r)e^{\iota k\varphi}, & r\leq 1,\\ 0, & r>1,\end{array}\right. \label{eq:prolatesdef}
\end{align}

where the range of $k,q$ is determined by Eq.~(8) in \cite{landa2017steerable}, the ${R_{k,q}}$ are a family of real, one-dimensional functions and the ${\alpha_{k,q}}$ are  scaling factors which will be defined in the next section.
The PSWFs are orthogonal on the unit disk.

For $\Bell \in [-(P-1), P-1]^2$, let us define 
\begin{equation} \label{eq:PSWF_expansion}
\mathfrak{a}^2[k, q]  =
\sum_{\Bell}a^2_\II[\Bell]\overline{\psi_{k,q}[\Bell]}, \end{equation}
where $\psi_{k,q}[\Bell] := \psi_{k,q}(\Bell/(P-1))$ is a discretization of the PSWFs. 
Knowledge of these coefficients is essentially equivalent to knowledge of the second-order correlations owing to the following approximate identity:
\begin{align}
a^2_\II[\Bell] & \approx \sum_{k,q}\mathfrak{a}^2[k,q]\psi_{k,q}[\Bell].
\label{eq:approxprolatesdiscrete}
\end{align}
This holds because the continuous PSWFs form an orthogonal basis, and their discretized counterparts are (empirically) almost orthogonal. As a result, for our purposes, the pair of equations above provides a basis expansion for the autocorrelations.

@@

Finally, we treat the bias terms:
\begin{align*}
&\mk b_3[k_1,q_1; k_2,q_2] = \gamma\sigma^2\langle a^1_{I_{\omega}} \rangle_\omega \delta[k_1 - k_2]\\&\times \left[\delta[q_1-q_2] + \delta[k_1]\frac{1}{2\pi}(\alpha_{0,q_1} + \alpha_{0,q_2})R_{0,q_1}(0)R_{0,q_2}(0)\right].
\end{align*}
%
Thus,
\begin{align*}
b_3[\Bell_1, \Bell_2] & = \gamma\sigma^2\langle a^1_{I_{\omega}}\rangle_{\omega}\Big(\delta[\Bell_1 - \Bell_2] +
\delta[\Bell_1] + \delta[\Bell_2]\Big) \\
& \approx \sum_{\substack{k_1, q_1\\ k_2, q_2}} \mk 
b_3[k_1,q_1; k_2, q_2]\psi_{k_1,q_1}[\Bell_1]\overline{\psi_{k_2,q_2}[\Bell_2]}.
\end{align*}

\section{Connection to volume}

@@

Using the 2-D PSWFs, we can express each projection as 
\[ \widehat I_{\omega}(ck,\theta) = \sum_{N,n}b_{N,n}(\omega)\psi_{N,n}(k,\theta),\]
where
\[\begin{aligned} b_{N,n}(\omega) &= \frac{4}{\sqrt{2\pi}|\alpha_{N,n}|^2}\int_0^{2\pi}\int_0^1\widehat I_{\omega}(ck,\theta)R_{N,n}(k)e^{-\iota N\theta}k\, dk\, d\theta,\\
&= \sum_{\ell,m,m',s}\tamir_{\ell,m,s}\left[\frac{2\sqrt{2\pi}}{\alpha_{N,n}}Y_{\ell}^{m'}(\pi/2,0)\right]D_{m',m}^{\ell}(\omega)\\&\times \left(\int_0^1j_{\ell,s}(k)R_{N,n}(k)k\, dk\right)\left(\frac{1}{2\pi}\int_0^{2\pi}e^{\iota(m'-N)\theta}\, d\theta\right),\\
&= \sum_{\vert N\vert\leq \ell}\sum_{m,s}\tamir_{\ell,m,s}D_{N,m}^{\ell}(\omega)\beta_{\ell,s;N,n},\end{aligned}\]
where the coefficients
\begin{equation} \label{eq:beta}
\beta_{\ell,s;N,n} := \left\{\begin{array}{ll} \frac{2\sqrt{2\pi}}{\alpha_{N,n}}Y_{\ell}^{N}(\pi/2,0)\int_0^1j_{\ell,s}(k)R_{N,n}(k)k\, dk, & \ell\geq |N|\\ 0, & \ell<|N|\end{array}\right.,
\end{equation}
can be precomputed.
Since the PSWFs are eigenfunctions of the truncated Fourier transform~\cite{landa2017steerable} and hence satisfy 
\begin{equation}
\label{eq:PSWF_defn_eq}
\alpha_{N,n}\psi_{N,n}(\mb k) = \int\limits_{||\mb r||_2\leq 1} \psi_{N,n}(\mb r)e^{\iota c(\mb r\cdot\mb k)}d\mb r,
\end{equation} 
we can now express the projection in real space as
\begin{align} \label{eq:proj_expansion}
I_{\omega}(r,\varphi) &= \sum_{N,n}\widehat{\alpha}_{N,n}b_{N,n}(\omega)\psi_{N,n}(r,\varphi)\nonumber \\
&= \sum_{\ell=0}^L\sum_{N,m=-\ell}^{\ell}\sum_{n=0}^{n_{\text{max}}(N)}\sum_{s=1}^{S(\ell)}\tamir_{\ell,m,s}\widehat\beta_{\ell,s;N,n}D_{N,m}^{\ell}(\omega)\psi_{N,n}(r,\varphi)
\end{align}
where $n_{max}(N)$ is chosen according to Eq.~(8) in \cite{landa2017steerable}, $\alpha_{N,n}$ is the eigenvalue corresponding to the $(N,n)$th PSWF, $\widehat{\alpha}_{N,n} = (c/2\pi)^2\alpha_{N,n}$, and $\widehat\beta_{\ell,s;N,n}=\widehat\alpha_{N,n}\beta_{\ell,s;N,n}$.

@@

%Since the Fourier transformed volume is given by
%\[ \widehat V(k,\theta,\varphi) =
%\sum_{\ell,m,s}a_{\ell,m,s}Y_{\ell}^m(\theta,\varphi)j_{\ell,s}(k),\]
Since $j_{\ell,s}(0) = 0$ unless $\ell=0$, and since
$Y_{0,0}(\theta,\varphi) = \frac{1}{\sqrt{4\pi}}$, we conclude from~\eqref{eq:volume_expansion_app} that
\begin{equation} \label{eq:ac_1_prolates}
\mathfrak{a}_x^1 = \langle
a^1_{I_{\omega}}\rangle_{\omega}  = \widehat  \phi(\mb 0) = \frac{1}{\sqrt{4\pi}}\sum_s\tamir_{0,0,s}j_{0,s}(0).
\end{equation}
